---
title: "GAM_MHD"
author: "Christen"
date: "March 25, 2020"
output: word_document
---

GAM Work with Monarch Health Data

*Research Question*
How do the spatial dynamics of OE infection prevalence change within year and between years in the sedentary population of monarch butterflies in the SE U.S.?

*Dataset*
```{r}
data <- read.csv("MHD_Full.csv",  colClasses = c("Pop_density_CAT" = "factor", "Hardiness2" = "factor", "Infection_Severity" = "factor", "Month"= "factor", "Year"= "factor"))
```

*Response Variable*
This is a binomial variable that assigns a 0 to samples where the infection level was below a 3. A 1 is assigned to infection levels above a 3#
```{r}
y <- data$Infection_Severity
```

*Random Effect*
```{r}
rand<-data$Volunteer.x
```

*Fixed Effects*
```{r}
Human_PopDens<- data$Pop_density_CAT
Month <- data$Month 
Winter_Year<- data$Season 
```

*Model Construction*
```{r}
require(mgcv)

# Fit the model
gam_mod <- gam(Infection_Severity ~ s(Human_PopDens, k=2, bs = "fs") + s(Winter_Year, k=2, bs = "fs") + s(Month, k=2, bs = "fs"), family = binomial, data = data)

# Plot the results
plot(gam_mod, residuals = TRUE, pch = 1)

summary(gam_mod)
```
```

