---
title: "GAM_MHD"
author: "Christen"
date: "March 25, 2020"
output:
  html_document:
    df_print: paged
---

GAM Work with Monarch Health Data



*Overall Research Question 1*
How do the spatial dynamics of OE infection prevalence change within year and between years in the sedentary population of monarch butterflies in the SE U.S.?




*Why use a GAM?*
We do not expect the patterns in disease prevalence as described by our predictor variables to be linear. For example, infection prevalence may experience increases during periods of warm weather such as in late fall and early spring, while mid winter low temperatures may result in a decline in infection prevalence as population density and therefore transmission potential declines. Graphs of the dataset are provided below for visual inspection of the linear or non-linear trends.  

*Dataset*
```{r}
wdata <- read.csv("./data/MHD_WD.csv", colClasses = c("Pop_density_CAT" = "factor", "Hardiness2" = "factor", "Infection_Severity" = "factor", "Month"= "factor", "Year"= "factor"))

data<- read.csv("./data/YR_MHD.csv")
```


*Exploration of the Data through Graphing*
Graph the relationship between infection and day of year to visually inspect the relationship for linearity. 
From this figure it is hard to discern any clear pattern. There seems to be a mostly random distribution of infection prevalence across all year (by day of year). 
```{r}
#Convert variables into different forms for graphing purposes
data$Day_Year<-as.factor(data$sDay_Year)
data$Infection_Severity<-as.numeric(data$Infection_Severity)

library(dplyr)
#Calculate the probability of infection on each day of year 
 library(plyr)
r2<-ddply(data,.(sDay_Year), summarize, mean=mean(Infection_Severity))
r3<- (r2$mean)-1
r2$r3 <- r3

library(ggplot2)
ggplot(data=r2, aes(x=as.numeric(sDay_Year), r3))+geom_point() + geom_smooth()

```

Graph the relationship between infection and year to visually inspect the relationship for linearity. 

```{r}
#Convert variables into different forms for graphing purposes
##data$Season<-revalue(data$Season, c("11_12"="1", "12_13"="2", "13_14"="3", "14_15"="4", "15_16"="5", "16_17"="6", "17_18"="7"))

class(data$Season)

data$Infection_Severity<-as.numeric(data$Infection_Severity)


#Calculate the probability of infection on each day of year 
 library(plyr)
s1<-ddply(data,.(Season), summarize, mean=mean(Infection_Severity))
s2<- (s1$mean)-1
s1$s2 <- s2

#Plot 
plot(as.numeric(s1$Season), s1$
       s2, xlab = "Year", ylab = "Probability of Infection", type='o')
```


*Analyzing the Change in Infection Prevalence Over Time*
  
  *Hypothesis 1: Within Year*
 Within a single winter season/year OE infection prevalence may increase with day of year as reproductive activity and increased monarch density lead to higher rates of transmission. 

  *Target output*: Create a model that analyzes how OE prevalence changes over the course of each individual season/year. 
  Create:
  1) 7 seperate figures, one for each season/year
  
 
  
  *Hypothesis 2: Between Years*
  If the population of winterbreeding monarchs in the southeastern US is growing, we expect that OE infection prevalence may increase with year as increased monarch density leads to higher rates of transmission.
  
  *Target output*: Create a model that analyzes how OE prevalance changes over time (years). 
Create: 
  1) One figure showing the change in infection over all 7 years of data 
  

*Model Construction*
This first model is designed to characterize temporal infection patterns in the southeast by including month, year and an interaction of these two terms in a single model. 

Details on the model:
1. Why "ti"?
  Tensor product smooths are used to model interactions between variables that have different natural scales. They multiply the smooth terms of one variable by the factor levels of the other. 

2. Why "by"?
  "By" variables are used for constructing "varying coefficient models" and for letting smooths interact with parametric terms. In this case our smooth is the Day_Year variable and the Season term is the parametric term (and a factor). 
The interaction between two parametric terms would use the : or * connotation, but smooth interactions cannot. "By" generates an indicator vector for each level of a factor  unless the factor is ordered. If it is ordered then a different smooth is generated for each fator level (except the first level). 


3. Why k=6? 
The term k is also referred to as "knots". These are the natural breaks in the basis functions that make up the "basis" or "smooth function" or "spline". Where the knots break along the spline is usually defined by quantiles. K is usually the max number of degrees of freedom allowed for a smooth term in the model. K in the models will always =  (max degrees of freedom - 1). In this case 

4. Why "family=binomial"? 
The response variable is either 0 (uninfected) or 1 (infected), this we have 2 curves for our response variable distribution and this requires that the model be fit with a binomial family distribution. 

5. Why is "Season" parametric?


6. Why is Volunteer "bs=re"? 
In order to include Volunteer as a random effect, we tell the model to use the random effect basis function to model this variable. This method only applies to mcgv, as a different method is used for gamm. 

7. Why REML? 

8. Should I be using a different basis function? Cubic? 

```{r}
require(mgcv)

#Ensure variables are of the correct type for model construction 
data$sDay_Year<-as.numeric(data$sDay_Year)
data$Season<-as.factor(data$Season)

data$Infection_Severity<-as.factor(data$Infection_Severity)

# Fit the model

gam_mod <- gam(Infection_Severity ~  Season + ti(sDay_Year, by = Season, k =6) + s(Volunteer.x, bs="re"), family = binomial, method = "REML", data = data)

#Code below for comparing models# 

###Ccomparing Season in the model with and without a smooth###

data$Season<-as.numeric(data$Season)


gam_modSS <- gam(Infection_Severity ~  s(Season, k= 6) + ti(sDay_Year, Season, k= 6) + s(Volunteer.x, bs="re"), family = binomial, method = "REML", data = data)

summary(gam_mod)
#gam_mod2 <- gam(Infection_Severity ~  Season + s(sDay_Year) + ti(sDay_Year, by = Season, k =6) + s(Volunteer.x, bs="re"), family = binomial, method = "REML", data = data)

#AIC(gam_mod, gam_mod2)
```
*Summarize output of model*

```{r}

#For table 
summary.gam(gam_mod) 

mgcv:summary(gam_mod)

class(gam_mod)

```
Key points from model output: 

The model output summary tells us that the additive temporal variables explain 36.7% of the deviance.We also can convert the outputs from the model to actual probabilities because the GAM model uses a log-odds scale for estimating outputs. After converting the intercept estimate below we find that the model predicts a 63.8% chance of infection overall. 

```{r}
plogis(0.5489)
```


*Checking the Model*
```{r}
#gam.check(gam_mod)

#gam.check does not seem to be the best method for visualizing logistic models

library(arm)

binnedplot(fitted(gam_mod), 
           residuals(gam_modSS, type = "response"), 
           nclass = NULL, 
           xlab = "Expected Values", 
           ylab = "Average residual", 
           main = "Binned residual plot", 
           cex.pts = 0.8, 
           col.pts = 1, 
           col.int = "gray")
```


*Plot the results of the temporal model* 
Plots of the partial effects of the interaction fo Day_Year and Season for each level of Season. 

```{r}
#install.packages("mgcViz")
require(mgcViz)

#Partial effects plots with 95% confidence intervals
  #y axis = Mean sqaure difference between fitted values? If transformed then we get probability 

#Version1: Figures with probabilities along y axis and shifted to incorporate the intercept of each smooth
  #these can be interpreted as the probability of the outcome if all other variables were at their average value. The confidence interval are the range of uncertainty of the probability of the outcome for any values of the variable, holding other variables equal at their average value (this interpretation is fromseWithMean)

plot(gam_mod, pages = 1, trans = plogis,
     shift = coef(gam_mod)[1], seWithMean = TRUE, rug = TRUE, shade = TRUE, shade.col = "lightgreen", 
     col = "purple")

#Version2
b <- getViz(gam_mod, trans = plogis)

print(plot(b, allTerms = T))
```

Visualization of the interaction between Season and Day_Year in a 3D plot. 

```{r}

vis.gam(gam_mod, n.grid = 50, theta = 400, phi = 20, ylab = "Day of Year", xlab = "Season", zlab = "",
        ticktype = "detailed", color = "terrain", main = "Interaction between day of year and season")



##theta= rotates the view around horizontally 
##phi = rotates the view vertically 
##n.grid =  controls the number of grid boxes
##z scale = GAM model predictions 

```

Contour plot of Day_year and Season. 

```{r}
vis.gam(GM,  plot.type = "contour", ylab = "Day of Year", xlab = "Season", color = "terrain", main = "Interaction between day of year and season", theme(legend.position = "top"))
GM<-(gam_mod, trans=plogis)

#Determine what the red lines mean 

source("https://sebastiansauer.github.io/Rcode/logit2prob.R")

GM<-logit2prob(coef(gam_mod))

require(ggplot2)
require(dbplyr)


```
```{r}

```


Summary of plot results: 

*Spatio-temporal Model Example* 

We see a change in years 2, 5, 6, 7, of infection prevalence increasing over Day_Year. Where do we end up with the highest infection prevalence? 

```{r}

#Create the background map 
library(dplyr)
library(maps)

south_east <- subset(states, region %in% c("florida", "georgia", "louisiana"))


se_map = map_data("state", "florida", "georgia")
                  
                  "florida", "georgia", "louisiana","texas")%>%
  transmute(Longitude= long, Latitude = lat, order=order)

```


```{r}
predict_richess = expand.grid(
  Latitude= seq(min(data$Latitude), 
                max(data$Latitude),
                length=50),
  Longitude = seq(min(data$Longitude),
                  max(data$Longitude),
                  length=50),
  Season = seq(1,7,by=1), 
  sDay_Year = seq(1,136,by=10), 
  Volunteer.x=1
)


predict_richess$model_fit = predict(gam_mod,
                                    predict_richess,type = "response")
ggplot(aes(Longitude, Latitude, fill= model_fit),
       data=predict_richess)+
  geom_tile()+
  facet_wrap(~Season,nrow=2)+
  scale_fill_viridis("# of species")+
  theme_bw(10)



predict_richess$model_fit = predict(gam_mod,
                                    predict_richess,type = "response")
predict_richess$model_change =predict(gam_mod,
                                      predict_richess%>%mutate(Season=Season+1),
                                      type = "response") -
  predict_richess$model_fit 


ggplot(aes(Longitude, Latitude, fill= model_fit),
       data=predict_richess)+
  geom_tile()+
  facet_wrap(~Season,nrow=2)+
  theme_bw(10)


gam_mod3 <- gam(Infection_Severity ~  ti(Latitude, Longitude) + Season + ti(sDay_Year, by = Season, k =6) + s(Volunteer.x, bs="re"), family = binomial, method = "REML", data = data)



```




*Research Question 2*
How do the listed exogenous variables impact the spatial dynamics of OE infection prevalence change within year and between years in the sedentary population of monarch butterflies in the SE U.S.?
  1. Human population density 
  2. Monarch larval density 
  3. Temperature 
  4. Precipitation
  

*Environmental Variables Model* 
In the Brown et al. paper, they used the GAMs in the previous step to determine the end of season infection prevalence for each year. They then looked for "bivariate correlations" between these infection prevalences and different environmental variables. 

Can I build a GAM to look at how infection varies with temperature, monarch density and other variables within and between years?


```{r}
#I could use the spatio temporal model to determine where OE is most prevalent. Then use these models to see why it is so prevalent in these locations. 

#Perhaps the question should be more focused - instead "which variables correlate with infection" it should be "which variables best predict the prevalnce of OE during a specific time (like in March when migrants return). THen we could use outputs from the temporal models (infection prevalence in march) as response variables"

Egam_mod <- gam(Infection_Severity ~  Season + ti(sDay_Year, by = Season, k =6) + s(EcoRegionIII) + Pop_density_CAT + s(DP10) + s(TMIN) + s(Volunteer.x, bs="re"), family = binomial, method = "REML", data = data)

#gam_mod <- gam(Infection_Severity ~  Season + ti(sDay_Year, by = Season, k =6) + s(Volunteer.x, bs="re"), family = binomial, method = "REML", data = data)

summary(Egam_mod)

```


*Research Question 3*
How do the observed patterns in infection dynamics differ by sex?
