---
title: "MHDWeatherData_Markdown"
author: "Christen"
date: "July 16, 2020"
output: html_document
---

Install the needed packages. 
```{r}
#When it asked you, did you choose binary or source? What if you do install.packages("XML", type="binary"). You are probably fine with the pre-compiled binary version. - MrFlick
install.packages("XML", type="binary")
install.packages("rnoaa")

library("XML")
library("rnoaa")
library("dplyr") 
```

Get through link the noaa code and save key. 
```{r}
options(noaakey = "mydUSdPqZMAjEcGHFedjxIeJWQBiBZjj")
```

Load a dataframe with the sample observations 
```{r}
data<- read.csv("./data/YR_MHD.csv")

```

Call the daily weather station data 
```{r}
library("rnoaa")
station_data <-ghcnd_stations() 
```

Create dataframe with just needed info to retrieve weather stations, removing duplicates with unique 
```{r}
lat_lon_df <- select(data,
                     City,
                     Latitude,
                     Longitude) %>% unique() %>%
  ungroup() %>%
  rename(id = City) %>%
   rename(latitude=Latitude) %>%
   rename(longitude =Longitude) %>%
  mutate(id = factor(id))

lat_lon_df = as.data.frame(lat_lon_df)

```

Find stations near the data points with the data needed 
```{r}
#Limit  controls the numer of stations pulled per site
MHDstations <- meteo_nearby_stations(lat_lon_df,
                                       station_data = station_data,
                                       limit = 5,
                                       year_min = 2011, year_max = 2018,
                                       var = ("all"))

MHDstations <- unique(bind_rows(MHDstations) %>% select(- distance))

```

Pull the actual weather data from the selected weather stations 
```{r}
monitors<-meteo_pull_monitors(monitors = MHDstations$id,
                            date_min = "2011-01-01", 
                             date_max = "2018-12-31", 
                                 var = "all")

save(monitors, file="monitors.Rdata")
```


Use the rnoaa package to get the coverage data for each weather station. You can then plot this data to visually see which stations have data. 
```{r}
#First use the meteo_coverage function to get the coverage per variable
obs_covr <- meteo_coverage(monitors)

library("ggplot2")
autoplot(obs_covr)
```

Filter the data by coverage 

```{r}

#Now convert the meteo object to a dataframe
library(tidyverse)
MonitorCoverage<-obs_covr %>%
  reduce(inner_join, by = "id")
MonitorCoverage<-as.data.frame(MonitorCoverage)

#Now filter out any stations with less than 80% coverage 
GoodMonitors<-MonitorCoverage %>% dplyr::filter(tmin.x >= 0.90)

```

Merge the good weather files with the original station data to get lat and long of each station 
```{r}
GoodMonitorsMerged=merge(GoodMonitors,MHDstations,by="id")



```




```{r}

#Prepare to combine WD file and MHD file based on year, month and state 
wd.date.get.year<-function(wd.date)
{
  as.numeric(strsplit(as.character(wd.date),split="-")[[1]][1])
}

wd.date.get.month<-function(wd.date)
{
  as.numeric(strsplit(as.character(wd.date),split="-")[[1]][2])
}

wd.date.get.day<-function(wd.date)
{
  as.numeric(strsplit(as.character(wd.date),split="-")[[1]][3])
}

wd.get.state<-function(wd.station.name)
{
  s<-strsplit(as.character(wd.station.name),split=" ")[[1]]
  s[length(s)-1]
}
GoodMonitorsMerged$Year<-sapply(GoodMonitorsMerged$date, wd.date.get.year)
GoodMonitorsMerged$Month<-sapply(GoodMonitorsMerged$date, wd.date.get.month)
GoodMonitorsMerged$Day<-sapply(GoodMonitorsMerged$date, wd.date.get.day)

write.csv(GoodMonitorsMerged, "GoodMonitorsMerged.csv", row.names = FALSE)
```


```{r}
#Rename columns in the data file to prepare for matching with the GoodMOnitorsMerged files
data <- data %>% rename( Year = Year.Sampled)
data <- data %>% rename( Month = Month.Sampled) 
data <- data %>% rename( Day = Day.Sampled) 
data <- data %>% rename( latitude = Latitude) 
data <- data %>% rename( longitude = Longitude) 


#After the preliminary sorting, find the closest weather station
find.closest<-function(point, stns)

{
  dist<-c()
  for(j in 1:nrow(stns))
  {
    dist[j]<-raster::pointDistance(point, c(stns$longitude[j],stns$latitude[j]),lonlat=TRUE)
  }
  which.min(dist)
  
}

stns<-data.frame()
for(i in 1:nrow(data))
{
  #for each obs in mh data, get (filter) all weather stns that are same year, month, state
  tmp<-filter(GoodMonitorsMerged, Year==data$Year[i] & Month==data$Month[i] & Day==data$Day[i]) 
        
  #find the closest station
  stns<-rbind(stns,tmp[find.closest(c(data$longitude[i],data$latitude[i]),tmp),])
  }



  
stns<-stns %>% transmute(Station=id,Station_Name=name,STN.LAT=latitude,STN.LON=longitude, 
                   TMIN=tmin.y,PRCP=prcp.y,PRCP_Coverage=prcp.x, TMIN_Coverage=tmin.x, TMAX= tmax.y, TAVG = tavg.y)

#Merge together the data file and the weather data 
weather_and_MHD<-cbind(data,stns)
#summary(weather_and_MHD)

#data[nrow(data) + 1,] = NA 


```

Edit reading in TMIN to reflect nearest 10th 
```{r}
weather_and_MHD$TMIN2 <-  weather_and_MHD$TMIN/10
weather_and_MHD$TMAX2 <-  weather_and_MHD$TMAX/10
weather_and_MHD$TAVG2 <-  weather_and_MHD$TAVG/10

#Editing of weather data into smaller data frame  
df = subset(DailyWeath, select = c(id, start_date, end_date, total_obs, tmin.x, tmax.y, tmin.y, date, PRCP= prcp.y))
df$TMIN <-  df$tmin.y/10
df$TMAX <-  df$tmax.y/10
df$TMIN_cov <-  df$tmin.x
df$PRCP <-  df$prcp.y
df$Date <-  df$date
df = subset(df, select = -c(tmin.x, tmax.y, tmin.y, date, prcp.y))

#Drop out unneeded columns 
df = subset(weather_and_MHD, select = -c(Method, Latitude_Zip_based, Longitude_Zip_based, TMIN, TMAX, TAVG, PRCP_Coverage, City.State) )


#Merge in Hardiness Zone data 
YRweather_hardi_MHD<-merge(Hardiness_Zipcode_Raw, weather_and_MHD, by.x="zipcode", by.y="Zipcode")
```

Write a csv file with the combined weather data and monarch health data 
```{r}
write.csv(YRweather_MHD, "YRweather_MHD.csv", row.names = FALSE)


write.csv(YRweather_hardi_MHD, "YRweather_hardi_MHD.csv", row.names = FALSE)

```

